Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-26 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
trimming text, adjusting (trying to) fix plot positions

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that it should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to each theory's fundamental length-scale $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
the Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final results using
$\varepsilon = 1/2$.
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %
We say that a bound has been placed on $\ell_{\rm th}$, if most of the PDF $P(\ell_{\rm th} | d)$ support is in the interval
$[0, \, \Lambda_{\rm EFT}(1/2, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated with the marginalized posterior distribution $P(\ell_{\rm th} | d)$, namely
%
\begin{equation}
P(\ell_{\rm th} \leqslant \ell^{\rm max}_{\rm th} | d) = \int_{0}^{\,\ell^{\rm max}_{\rm th}} \dd \ell'\, P(\ell' | d).
\label{eq:def_cdf}
\end{equation}
%
For instance, we require that for a bound at 90\% credible level to be placed on $\ell_{\rm th}$ that
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell_{\rm th} \leqslant \Lambda_{\rm EFT} | d) \leqslant 0.9,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
where we let $\ell^{\rm max}_{\rm th} = \Lambda_{\rm EFT}$ in Eq.~\eqref{eq:def_cdf}, and
likewise for other credibility percentiles.

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function on coupling constant $\ell_{\rm EdGB}$ in EdGB gravity
for GW150914 (top panel), GW200129 (middle panel)
and from combining results (bottom panel).
%
In all panels, different line colors correspond to the inclusion ($N_{\rm max} = 1$) or not ($N_{\rm max} = 0$)
of the linear-in-spin QNM correction.
%
The joint posteriors are shown for illustrative purposes only. As we explain in Fig.~\ref{fig:edgb_cdf} and
in the main text, our analysis of these events fails to satisfy the EFT bound~\eqref{eq:eft_bound}. We mark
with vertical lines theh $90\%$ upper credible intervals.}
\label{fig:edgb_exec_sum}
\end{figure}

@@

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The CDF evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon,\gm)$ for EdGB gravity as a function of the parameter
$\varepsilon$ for both (dashed curves) $\gm = m_2$, the secondary's source
mass, and (solid curves) $\gm = M_{\rm f}$, the remnant's source mass, without
(black curves) and with (purple curves) linear in spin QNM corrections. The
horizontal lines mark the $90\%$ credible levels. Having set the maximum value
of $\varepsilon$ to be 1/2, we see that in no situation the curves pass
through the 0.90 lines. This means that no bound on $\ell_{\rm EdGB}$ can be placed
with the events we analyzed.
}
\label{fig:edgb_cdf}
\end{figure}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event &  EFT    & ParSpec & Constraint      \\
              &       &  bound? & bound?  & ($\gm = M_{f}$) \\
\hline
  & GW150914 & No  & Yes & -- \\
0 & GW200129 & No  & Yes & -- \\
  & Combined & --  & Yes & -- \\
\hline
  & GW150914 & No  & Yes & -- \\
1 & GW200129 & No  & Yes & -- \\
  & Combined & --  & Yes & -- \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results for EdGB gravity for GW150914, GW200129, and
combined events using $\gm = M_{\rm f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
We find that we cannot place any constraint on $\ell_{\rm EdGB}$ with our waveform model from either GW event.
}
\label{tab:summary_edgb}
\end{table}

@@

The QNMs of slowly-rotating BHs in both cubic and quartic EFT of GR where calculated in Ref.~\cite{Cano:2021myl}.
%
For the cubic EFT, we use their Eq.~(67), particularizing \ab{it} to $\lame = \lamo = 1$.
%
We further linearize that equation in $\chi$ and \sout{particularize} \ab{consider} $m=2$.
%
This process, shows that the mode labeled ``-'' is the least-damped \ab{mode}, and it is the one we use.
%
Direct comparison with Eqs.~\eqref{eq:qnm_nongr_iso} results in $p_{\rm cEFT} = 4$,
%
\begin{align}
    \delta \omega_{\rm cEFT}^{(0)} = - 0.5813, \quad \delta \tau_{\rm cEFT}^{(0)} = 2.6469,
    \nonumber \\
    \delta \omega_{\rm cEFT}^{(1)} = -3.8620, \quad \delta \tau_{\rm cEFT}^{(1)} = 265.12.
\end{align}

We proceed in the same away for the quartic EFT. Here we use Eq.~(68) (with
$\epsilon_{1} = 1$ and $\epsilon_{2} = 0$) and Eq.~(70) of
Ref.~\cite{Cano:2021myl},
%
In this case we find that both modes (``-'' and ``+'' in the notation of Ref.~\cite{Cano:2020cao}) make the
fundamental QNM damp fast, and we choose the ``-'' mode which decreases less the damping time. \abcomm{Not clear.
Decreases less with respect to what? Do you mean it has a smaller damping time?}
%
\ab{This time we} then find \sout{this time} that $p_{\rm qEFT} = 6$,
%
\begin{align}
    \delta \omega_{\rm qEFT}^{(0)} = -0.2114, \quad \delta \tau_{\rm qEFT}^{(0)} = -0.6070,
    \nonumber \\
    \delta \omega_{\rm qEFT}^{(1)} = -1.5263, \quad \delta \tau_{\rm qEFT}^{(1)} = 171.35.
\end{align}
%
As with the previous theories, the larger values of some of these parameters are compensated
by the assumptions of weak coupling and small spin used to calculate the QNM frequencies.
\
\section{The estimation of  intrinsic binary parameters in General Relativity and modified theories of gravity}
\label{app:big_contour}

>>>>

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that it should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to each theory's fundamental length-scale $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final
results using $\varepsilon = 1/2$, \hs{but we stress that there is no
fundamental justification for using this cut off.}
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %

Under these assumptions, we will say that a bound has been placed on $\ell_{\rm th}$, if most of the PDF $P(\ell_{\rm th} | d)$ support is in the interval
$[0, \, \Lambda_{\rm EFT}(1/2, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated with the marginalized posterior distribution $P(\ell_{\rm th} | d)$, namely
%
\begin{equation}
P(\ell_{\rm th} \leqslant \ell^{\rm max}_{\rm th} | d) = \int_{0}^{\,\ell^{\rm max}_{\rm th}} \dd \ell'\, P(\ell' | d).
\label{eq:def_cdf}
\end{equation}
%
For instance, we require that for a bound at 90\% credible level to be placed on $\ell_{\rm th}$ that
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell_{\rm th} \leqslant \Lambda_{\rm EFT} | d) \leqslant 0.9,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
where we let $\ell^{\rm max}_{\rm th} = \Lambda_{\rm EFT}$ in Eq.~\eqref{eq:def_cdf}, and
likewise for other credibility percentiles.

@@

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{figs/edgb_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function on coupling constant $\ell_{\rm EdGB}$ in EdGB gravity
for GW150914 (top panel), GW200129 (middle panel)
and from combining results (bottom panel).
%
In all panels, different line colors correspond to the inclusion ($N_{\rm max} = 1$) or not ($N_{\rm max} = 0$)
of the linear-in-spin QNM correction.
%
The joint posteriors are shown for illustrative purposes only. As we explain in Fig.~\ref{fig:edgb_cdf} and
in the main text, our analysis of these events fails to satisfy the EFT bound~\eqref{eq:eft_bound}. We mark
with vertical lines theh $90\%$ upper credible intervals.}
\label{fig:edgb_exec_sum}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{figs/edgb_cdf_varying_threshold.pdf}
\caption{(Color online).~The CDF evaluated at the
cut-off $\Lambda_{\rm EFT}(\varepsilon,\gm)$ for EdGB gravity as a function of the parameter
$\varepsilon$ for both (dashed curves) $\gm = m_2$, the secondary's source
mass, and (solid curves) $\gm = M_{\rm f}$, the remnant's source mass, without
(black curves) and with (purple curves) linear in spin QNM corrections. The
horizontal lines mark the $90\%$ credible levels. Having set the maximum value
of $\varepsilon$ to be 1/2, we see that in no situation the curves pass
through the 0.90 lines. This means that no bound on $\ell_{\rm EdGB}$ can be placed
with the events we analyzed.
}
\label{fig:edgb_cdf}
\end{figure}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
$N_{\rm max}$ & Event &  EFT    & ParSpec & Constraint      \\
              &       &  bound? & bound?  & ($\gm = M_{f}$) \\
\hline
  & GW150914 & No  & Yes & -- \\
0 & GW200129 & No  & Yes & -- \\
  & Combined & --  & Yes & -- \\
\hline
  & GW150914 & No  & Yes & -- \\
1 & GW200129 & No  & Yes & -- \\
  & Combined & --  & Yes & -- \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results for EdGB gravity for GW150914, GW200129, and
combined events using $\gm = M_{\rm f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
We find that we cannot place any bound on $\ell_{\rm EdGB}$ with our waveform model from either GW event.
}
\label{tab:summary_edgb}
\end{table}

@@

The QNMs of slowly-rotating BHs in both cubic and quartic EFT of GR where calculated in Ref.~\cite{Cano:2021myl}.
%
For the cubic EFT, we use their Eq.~(67), particularizing it to $\lame = \lamo = 1$.
%
We further linearize that equation in $\chi$ and consider the harmonic $m=2$.
%
This process, shows that the mode labeled ``-'' is the least-damped mode, and it is the one we use.
%
Direct comparison with Eqs.~\eqref{eq:qnm_nongr_iso} results in $p_{\rm cEFT} = 4$,
%
\begin{align}
    \delta \omega_{\rm cEFT}^{(0)} = - 0.5813, \quad \delta \tau_{\rm cEFT}^{(0)} = 2.6469,
    \nonumber \\
    \delta \omega_{\rm cEFT}^{(1)} = -3.8620, \quad \delta \tau_{\rm cEFT}^{(1)} = 265.12.
\end{align}

We proceed in the same away for the quartic EFT. Here we use Eq.~(68) (with
$\epsilon_{1} = 1$ and $\epsilon_{2} = 0$) and Eq.~(70) of
Ref.~\cite{Cano:2021myl},
%
In this case we find that both modes (``-'' and ``+'' in the notation of Ref.~\cite{Cano:2020cao}) make the
fundamental QNM damp fast, and we choose the ``-'' mode which decreases less the damping time. \abcomm{Not clear.
Decreases less with respect to what? Do you mean it has a smaller damping time?}
%
This time we then find that $p_{\rm qEFT} = 6$,
%
\begin{align}
    \delta \omega_{\rm qEFT}^{(0)} = -0.2114, \quad \delta \tau_{\rm qEFT}^{(0)} = -0.6070,
    \nonumber \\
    \delta \omega_{\rm qEFT}^{(1)} = -1.5263, \quad \delta \tau_{\rm qEFT}^{(1)} = 171.35.
\end{align}
%
As with the previous theories, the larger values of some of these parameters are compensated
by the assumptions of weak coupling and small spin used to calculate the QNM frequencies.
\
\section{The estimation of  intrinsic binary parameters in General Relativity and modified theories of gravity}
\label{app:big_contour}
